### 2023-04-20 Experiment_2/analysis.Rmd

---
title: "Analysis experiment 2"
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-09"
output: 
  html_document: 
    keep_md: yes
---

```{r packages, message=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library(ggrepel)       # annotations in plots
```

```{r, message=FALSE}
d <- read_csv("./data/cleaned.csv")
```

### Some themes and functions

```{r, echo=FALSE}
# round numbers from models
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 3)

#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

## Data analysis

### Hypothesis 1

To assess the effect of independence on participants’ confidence about their guesses, we run a paired t-test.

```{r}
# get summary stats by number
d %>% 
  group_by(independence) %>% 
  summarise(across(c(confidence, competence),  
                   list(mean = mean, sd = sd)))
```

```{r}
# T-test
t <- t.test(confidence ~ independence, data = d, paired = TRUE)
t
```
The paired t-test shows a significant difference in confidence ratings between _independence_ and _dependence_. In the independence condition, participants were on average more confident by `r abs(t$estimate[[1]])` points (on a scale from 1 to 7).

For a robustness check, we run a mixed model regression analysis of `independence` on `confidence` ratings. The first model (`m1`) includes random intercepts for participants.

For reference, we start with standard, fixed-only effects regression models. 
```{r}
# for confidence 
# fixed linear regression models for reference
fixed_confidence <- lm(confidence ~ independence, data= d)
# model with random intercept per participant
mixed_confidence = lmer(confidence ~ independence + (1 |ID), data = d)

# for competence 
# fixed linear regression models for reference
fixed_competence <- lm(competence ~ independence, data= d)
# model with random intercept per participant
mixed_competence= lmer(competence ~ independence + (1 |ID), data = d)
# store in list
models = list("fixed" = fixed_confidence,
              "mixed" = mixed_confidence, 
              "fixed" = fixed_competence, 
              "mixed" = mixed_competence)
```

```{r}
modelsummary(models,
             stars = TRUE,
             title = 'Models estimating effect of convergence on accuracy and competence') %>%
    add_header_above(c(" " = 1, "Accuracy" = 2, "Competence" = 2))
```

The fixed-effect-only linear regression model yields the same estimate for the effect of `independence` as the paired t-test, with the difference that this is not statistically significant at the 5%-level anymore. In the mixed model, the main effect of `independence` remains the same, but is significant again, like in the paired t-test.

### Hypothesis 2

We will proceed just as with hypothesis 1, but for `competence`.

```{r}
# T-test
t <- t.test(competence ~ independence, data = d, paired = TRUE)
t
```
The paired t-test does not show a significant difference in confidence ratings between _no convergence_ and _convergence_.

## Visualization

### Main plot

```{r}
# make plots
confidence_plot <- ggplot(data=d, aes(x=confidence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.45)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1, 
                    name = "Independence: ",
                    limits = c("dep", "indep"), 
                    labels = c("discussion", "independence")) + 
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

competence_plot <- ggplot(data=d, aes(x=competence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.45)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1, 
                    name = "Independence: ",
                    limits = c("dep", "indep"), 
                    labels = c("discussion", "independence")) + 
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

figure <- ggarrange(confidence_plot, competence_plot, common.legend = T) +
  theme(legend.position="top") 
figure
```

### Other plots

Here's a visualization of the difference in `confidence` ratings between the _independence_ and _dependence_ conditions. 
```{r}
# create objects for means to include in graphics
d_means <- d %>%
  group_by(independence) %>%
  summarize(mean_confidence=mean(confidence))

# Vizualization
density_plot <- ggplot(data=d, aes(x=confidence, fill=independence)) +
geom_density(alpha=.4) 
density_plot

density_plot_smoothed <- ggplot(data=d, aes(x=confidence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Confidence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_fill_brewer(palette = "Dark2") +
  theme_pubr()
density_plot_smoothed

# prettier
density_plot_smoothed <- ggplot(data=d, aes(x=confidence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Confidence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
  geom_vline(data=d_means, aes(xintercept=mean_confidence, 
                                       color=independence),
             show.legend = FALSE,
             linetype="dashed", size=1.5) +
  geom_text_repel(data = d_means, 
            aes(x = mean_confidence, y = 0.3, 
                label =  paste("mean", round(mean_confidence, digits = 1),
                               sep = " = "))) +
  scale_fill_brewer(
    name = "Independence: ",
    labels = c("discussion", "independence"),
    palette = "Accent") + 
  scale_color_brewer(palette = "Accent") +
  guides(fill = guide_legend(title = NULL)) +
  theme_pubr() + 
  # larger axis titles
  theme(axis.title = element_text(size = 20)) +
  theme(legend.text = element_text(size = 15))
density_plot_smoothed

```

Here's a visualization of the difference in `competence` ratings between the _independence_ and _dependence_ conditions. 
```{r}
# create objects for means to include in graphics
d_means <- d %>%
  group_by(independence) %>%
  summarize(mean_competence=mean(competence))

# Vizualization
density_plot <- ggplot(data=d, aes(x=competence, fill=independence)) +
geom_density(alpha=.4) 
density_plot

# prettier
density_plot_smoothed <- ggplot(data=d, aes(x=competence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.5)) +
  geom_vline(data=d_means, aes(xintercept=mean_competence, 
                                       color=independence),
             show.legend = FALSE,
             linetype="dashed", size=1.5) +
  geom_text_repel(data = d_means, 
            aes(x = mean_competence, y = 0.47, 
                label =  paste("mean", round(mean_competence, digits = 1),
                               sep = " = "))) +
  scale_fill_brewer(
    name = "Independence: ",
    labels = c("discussion", "independence"),
    palette = "Accent") + 
  scale_color_brewer(palette = "Accent") +
  guides(fill = guide_legend(title = NULL)) +
  theme_pubr() + 
  # larger axis titles
  theme(axis.title = element_text(size = 20)) +
  theme(legend.text = element_text(size = 15))
density_plot_smoothed

```