### 2023-04-20 Experiment_4/old_model.Rmd

---
title: "Model"
output: 
  html_document: 
    keep_md: yes
date: "2023-04-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r, Message=FALSE}
library(tidyverse)
library(gghalves) # for plots
library(patchwork)
```

## Set plot theme
```{r}
#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

## Function for single population
```{r}
simulate_single_population <- function(population, sample, options) {

  # randomly draw competence levels
  data <- tibble(id = 1:population,
                 competence = runif(population, min = 1/options, max = 1))

  # randomly draw individual answers based on competence levels 
  data$answer <- data$competence %>% 
    purrr::map_chr(function(x){

      answer_options <- c("correct", paste0("false", 1:(options-1)))
      probabilities <- c(x, rep((1-x)/(options-1), options-1))

      answer = sample(answer_options, 
                      size = 1, 
                      prob = probabilities
      )
      return(answer)

    })

  # randomly assign samples in population
  data <- data %>% mutate(sample = rep(1:(population/sample), sample))

  # assign constellations build a variable identifying accurate responses
  data <- data %>% group_by(sample) %>% 
    mutate(unique_answers = n_distinct(answer),
           # Build a first version of constellation variable
           constellation = case_when(unique_answers == 3 ~ "dissensus", 
                                     unique_answers == 1 ~ "consensus",
                                     unique_answers == 2  ~ "majority"
           ), 
           # identify minority answers in majority constellations
           minority = ifelse(constellation == "majority", 
                             # negate duplicated answers
                             !(duplicated(answer) | duplicated(answer, 
                                                               # makes sure that the original value AND
                                                               # the duplicate get marked as duplicates
                                                               fromLast = TRUE)),
                             # for all other constellations, simply code NA,
                             NA
           ), 
           # modify the constellation variable to distinguish between minority 
           # and majority
           constellation = ifelse(is.na(minority) | minority == FALSE, 
                                  constellation, 
                                  "minority")
    ) %>% ungroup()

  # identify accurate responses
  data <- data %>%
    mutate(accurate = ifelse(answer == "correct", TRUE, FALSE))

  # compute the summary statistics by constellation
  results <- data %>% 
    group_by(constellation) %>% 
    summarize(average_competence = mean(competence), 
              average_accuracy = mean(accurate)) %>% 
    # store simulation information
    mutate(population = population, 
           sample = sample, 
           options = options)

  return(results)
}

```

## Function for various populations
```{r}
simulate_various_populations <- function(iterations,...) {

  # create data frame with model results for generated samples
  various_populations <- 1:iterations %>% 
    purrr::map_df(function(x){
      # this is essentially a for loop - do the following for each 
      # element in 1:iterations

      results <- simulate_single_population(...)

      # identify iteration
      results$iteration <- x

      # To keep track of progress
      if (x %% 50 == 0) {print(paste("iteration number ", x))}

      return(results)

    }) %>% 
    # store simulation information
    mutate(total_iterations = iterations)

  return(various_populations)
}
```

## Plot function
```{r}
plot_results <- function(data) {

  d <- data

  # make constellation a factor with the right levels
  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "majority", "consensus")

  simulation_info <- d %>% summarize(across(c(population, sample, options, total_iterations), mean)) %>% 
    pivot_longer(everything(), names_to = "parameter", values_to = "value") %>% 
    gridExtra::tableGrob(cols = NULL, rows = NULL)

  # plot for accuracy
  plot_accuracy <- ggplot(d,
                          aes(x = constellation, y = average_accuracy, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = "Convergence", y = "Accuracy") +
    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
    guides(fill = FALSE) +
    plot_theme + 
    theme(axis.text.x = element_text(angle = 20, hjust = 1))

  # plot for competence
  plot_competence <- ggplot(d,
                            aes(x = constellation, y = average_competence, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = "Convergence", y = "Competence") +
    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
    guides(fill = FALSE) +
    plot_theme + 
    theme(axis.text.x = element_text(angle = 20, hjust = 1))

  (plot_accuracy | plot_competence) / simulation_info
}
```

## Simulate

Note that in the current state, we cannot vary `sample`. The function needs modification first. The problem is that it gets complicated with more than three informants, in the case of a majority constellation. There might be equal majorities (e.g. 5 informants, three options: 2 agree on 'A', 2 on 'B', one picks 'A'), equal minorities (e.g. 3 agree on 'A', one picks 'B', another picks 'C'), and unequal majorities (e.g. 3 agree on 'A', 2 agree on 'B').

I have to think about it again.

```{r}
data_3 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 3)
data_10 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 10)
```

```{r}
plot_results(data_3)
plot_results(data_10)
```