### 2023-05-22 Experiment_4/analysis.Rmd
add experiment 6 and slight model updates

---
title: 
- "Analysis Experiment 4: Do inferences from convergence hold in a categorical choice setting?"
author: 
- "Hugo Mercier, Jan Pfänder" 
date: "2023-02-02"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = TRUE) 
```

```{r, message=FALSE}
library(tidyverse) # for everything
library(tidyr) # some additional, neat tidy reshaping functions
library(lme4) # for linear mixed models
library(lmerTest) # p-values for mixed models
library(broom) # for tidy outputs of regressions
library(broom.mixed) # for tidy outputs of linear mixed models
library(ggridges) # for plots
library(gghalves) # for plots
library(ggbeeswarm) # Special distribution-shaped point jittering
library(scales)  # for nice axis labels
library(modelsummary) # regression tables
library(patchwork) # combine plots
library(ggrepel) # fit labels in plots 
```

### Some themes and functions

```{r}
# round numbers from models
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 3)

#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

### Read data
```{r, message=FALSE}
# read data
d <- read_csv("./data/cleaned.csv")
```

### H1a: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.

We use a linear mixed effect model with random intercept and random slope per participant. Should this model yield convergence issues, we will use a model with random intercept only.

In all our models we treat `convergence` as a continuous variable. We will, however, include robustness checks where we treat convergence as a categorical variable, allowing to inspect difference between different levels.

```{r}
# model for accuracy
# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence + (1 + convergence | id), 
                                 data = d)
```

### H1b: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.

We will proceed in the same way for `competence` as we did for `accuracy` above.

```{r}
# model for competence
# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence + 
                           (1 + convergence | id), data = d)
```

```{r}
# combined results table
modelsummary(list("Accuracy" = model_accuracy , "Competence" = model_competence), 
             title = 'Models estimating effect of Convergence', 
             stars = TRUE)
```

# Robustness checks

In the models above, we treated convergence as a continuous variable. We run the same models outlined in the hypotheses section, but replacing `convergence` with `convergence_categorical` (where all levels are categories). This also allows us to inspect heterogeniety in differences between levels (with respect to the baseline, i.e. "opposing majority").

```{r}
# show levels of `convergence`
levels(as.factor(d$convergence_categorical))
```

```{r}
# model for accuracy
# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence_categorical + (1 + convergence_categorical | id), 
                                 data = d)
# model for competence
# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence_categorical + 
                           (1 + convergence_categorical | id), data = d)
```

```{r}
# combined results table
modelsummary(list("Accuracy" = model_accuracy , "Competence" = model_competence), 
             title = 'Models estimating effect of Convergence', 
             stars = TRUE)
```

### Plots

```{r}
# plot for accuracy
plot_accuracy <- ggplot(d,
       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r}
# plot for competence
plot_competence <- ggplot(d,
       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r}
plot_accuracy + plot_competence
```