### 2023-11-08 Experiment_4/model.Rmd

---
title: "Model for categorical choice scenarios (experiments 4 to 6)"
output: 
  html_document: 
    keep_md: yes
date: "2023-04-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, echo=FALSE}
# Load packages
library(tidyverse)
library(gghalves) # for plots
library(patchwork)
library(lme4) # for linear mixed models
library(lmerTest) # p-values for mixed models
library(broom) # for tidy outputs of regressions
library(broom.mixed) # for tidy outputs of linear mixed models
library(ggpubr) # for plots
```

```{r, echo=FALSE}
# set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold title
    plot.title = element_text(face = "bold", size = rel(0.7)),
    # Plain subtitle that is grey
    plot.subtitle = element_text(face = "plain", size = rel(1), color = "grey70"),
    # Bold legend titles
    legend.title = element_text(face = "bold"),
    # Bold axis titles
    axis.title = element_text(face = "bold"), 
    # Bold legend labels
    legend.text = element_text(face = "bold")
    )
```

```{r}
# ensure this script returns the same results on each run
set.seed(7643)
```

## Explaining the model

```{r}
# Imagine a situation with 3 choice options
options <- 3
# and a population of n = 999
population <- 999
# and we observe samples of 3
sample <- 3
```

### Define competence and its distribution

We define competence as the probability of making an accurate choice (among various options). Competence is continuous and ranges from `0` to `1`, where `0` corresponds to pure chance (`1/options`) and `1` corresponds to certainly making the correct choice.

In our code this measure is called `relative_competence`, while `competence` designates the actual probability of making an accurate choice, ranging from `1/options` to `1`). For example, in a 3-choice-options scenario, an individual with a probability of 1/3 (`competence == 0.333`) to pick the right answer has `relative_competence == 0`.

We assume that individuals (i) vary in competence and (ii) that the distribution of competence in the population varies, as can be modeled by a beta distribution. In the case below, we assume a uniform competence distribution (equal to a beta distribution of alpha = 1 and beta = 1).

```{r}
# 1: Distribution of competence
data <- data.frame(competence = runif(population, min = 1/options, max = 1))

# Create a relative measure of competence that allows to compare across
# different numbers of choice options. 
# This measure varies from 0 = random chance to 1 = definitely right. 
# We use min-max scaling, with min = 1/options and max = 1.
data <- data %>% 
  mutate(relative_competence = (competence - 1/options) / (1 - 1/options) )

# data generating function
ggplot() +
  stat_function(fun = dunif, args = list(min = 0, max = 1)) +
  labs(title = "Uniform distribution of competence from 0 to 1", x = "Competence", y = "P(Competence)") + 
  xlim(0, 1)

# generated population
ggplot(data, aes(x = relative_competence)) +
  geom_histogram() + 
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) + 
  annotate("text", x = 0.9, y = 100, label = paste0("n = ", population), color = "red") +
  labs(title = "(sampled) Population of competence drawn from \n uniform population distribution", 
       y = "Count")
```

### Generate a choice for each individual based on their competence.

```{r}
# 2: Draw individual answers based on competence levels 
data$answer <- data$competence %>% 
  purrr::map_chr(function(x){

    answer_options <- c("correct", paste0("false", 1:(options-1)))
    probabilities <- c(x, rep((1-x)/(options-1), options-1))

    answer = sample(answer_options, 
                    size = 1, 
                    prob = probabilities
                    )
  }
  )

data <- data %>% 
  mutate(correct = ifelse(str_detect(answer, "correct"), 
                                        TRUE, FALSE))

answers <- ggplot(data, aes(x = answer, fill = correct)) +
  geom_bar() +
  guides(fill = "none") +
  plot_theme

competence_by_answers <- ggplot(data, 
       aes(x = answer, y = relative_competence, fill = correct)) +
geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .8,
                     side = "r") +
  coord_flip() +
  guides(fill = "none") +
  plot_theme

answers + competence_by_answers
```

### Draw samples, assign categories & compare average outcomes

### Randomly draw samples from the population

```{r}
# 3: randomly assign samples in population
data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))
```

### Categorize constellations.

```{r}
# 4: identify constellations 
data <- data %>% 
  # identify how often a one type of answer occurs in one group
  group_by(sample_id, answer) %>% 
  mutate(n_answer_in_sample = n()) %>% 
  # assign constellations
  group_by(sample_id) %>% 
  mutate(unique_answers = n_distinct(answer),
         # Build a first version of constellation variable
         constellation = case_when(unique_answers == 1 ~ "consensus",
                                   unique_answers <=  sample ~ "majority-minority-dissensus"
         ),
         # identify majority answers in majority constellations
         majority = ifelse(constellation == "majority-minority-dissensus", 
                           # report whether occurrences of answer within a group
                           # are the majority within that group
                           n_answer_in_sample == max(n_answer_in_sample),
                           # for all other constellations, simply code NA,
                           NA
         ), 
         # identify dissensus answers in majority constellations
         dissensus = ifelse(constellation == "majority-minority-dissensus", 
                            # report whether occurrences of answer within a group
                            # are the minority within that group
                            min(n_answer_in_sample) == max(n_answer_in_sample),
                            # for all other constellations, simply code NA,
                            NA
         ), 
         # build final constellation variable using helper variables 
         constellation = case_when(
           dissensus == TRUE ~ "dissensus",
           majority == TRUE ~ "majority",
           majority == FALSE ~ "minority(ies)",
           .default = constellation)
  ) %>% ungroup()
```

### Compute average accuracy/competence by constellation.

```{r}
# 5: calculate accuracy and competence levels per constellation

# identify accurate responses
data <- data %>%
  mutate(accurate = ifelse(answer == "correct", TRUE, FALSE))

# compute the summary statistics by sample id
results <- data %>% 
  group_by(sample_id, constellation) %>% 
  summarize(average_competence = mean(competence), 
            average_relative_competence = mean(relative_competence),
            average_accuracy = mean(accurate)) %>% 
  # store simulation information
  mutate(population = population, 
         sample = sample, 
         options = options)
```

```{r}
ggplot(results,
       aes(x = constellation, y = average_accuracy, fill = constellation)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = "none") +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
  ylim(c(0, 1))
```

## Functions

### Data generating functions

#### Single population

```{r, echo=FALSE}
simulate_single_population <- function(population, sample, options, 
                                       distribution = "uniform", 
                                       alpha, beta,
                                       return = "by constellation") {
  # Check if population is not divisible by sample without remainder
  if (population %% sample != 0) {
    warning("Population is not divisible by sample without remainder.")
  }

  # if necessary, alter population size so that it is divisible by sample without rest
  n_possible_samples <- floor(population/sample)

  # Biggest possible population to pick
  possible_population = n_possible_samples * sample

  # Issue warning if different population size used
  if (population %% sample != 0) {
    warning(paste0("Chosen population size is ", possible_population))
  }
  # change population value (if divisible without remainder, it'll be the same)
  population = possible_population

  # 1: Distribution of competence

  if (distribution == "uniform") {
  # randomly draw competence levels
    data <- tibble(id = 1:population,
                   competence = runif(population, min = 1/options, max = 1)) %>% 
      # Create a relative measure of competence that allows to compare across
      # different numbers of choice options. 
      # This measure varies from 0 = random chance to 1 = definitely right. 
      # We use min-max scaling, with min = 1/options and max = 1.
      mutate(relative_competence = (competence - 1/options) / (1 - 1/options))
  }

  if (distribution == "beta") {
  # randomly draw competence levels
  data <- tibble(id = 1:population,
                 relative_competence = rbeta(population, shape1 = alpha, 
                                             shape2 = beta)) %>% 
    mutate(
      competence = relative_competence * (1 - 1/options) + 1/options
      )
  }

  # 2: Draw individual answers based on competence levels 
  data$answer <- data$competence %>% 
    purrr::map_chr(function(x){

      answer_options <- c("correct", paste0("false", 1:(options-1)))
      probabilities <- c(x, rep((1-x)/(options-1), options-1))

      answer = sample(answer_options, 
                      size = 1, 
                      prob = probabilities
      )
      return(answer)

    })

  # 3: randomly assign samples in population
  data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))

  # 4: identify constellations 
  data <- data %>% 
    # identify how often a one type of answer occurs in one group
    group_by(sample_id, answer) %>% 
  mutate(n_answer_in_sample = n()) %>% 
    # assign constellations
  group_by(sample_id) %>% 
    mutate(unique_answers = n_distinct(answer),
           # relative degree of majority
           n_sample = n_distinct(id),
           relative_majority = n_answer_in_sample / n_sample,
           # Build a first version of constellation variable
           constellation = case_when(unique_answers == 1 ~ "consensus",
                                     unique_answers <=  sample ~ "majority-minority-dissensus"
           ),
           # identify majority answers in majority constellations
           majority = ifelse(constellation == "majority-minority-dissensus", 
                             # report whether occurrences of answer within a group
                             # are the majority within that group
                             n_answer_in_sample == max(n_answer_in_sample),
                             # for all other constellations, simply code NA,
                             NA
           ), 
           # identify dissensus answers in majority constellations
           dissensus = ifelse(constellation == "majority-minority-dissensus", 
                             # report whether occurrences of answer within a group
                             # are the minority within that group
                             min(n_answer_in_sample) == max(n_answer_in_sample),
                             # for all other constellations, simply code NA,
                             NA
           ), 
           # build final constellation variable using helper variables 
           constellation = case_when(
             dissensus == TRUE ~ "dissensus",
             majority == TRUE ~ "majority",
             majority == FALSE ~ "minority(ies)",
             .default = constellation)
           ) %>% ungroup()

  # 5: calculate accuracy and competence levels per constellation

  # identify accurate responses
  data <- data %>%
    mutate(accurate = ifelse(answer == "correct", TRUE, FALSE))

  if(return == "by participant") {
    return(data)
  } else {

    if(return == "by degree of majority") {
      results <- data %>% 
        group_by(sample_id, relative_majority)
    }

    if(return == "by constellation") {
      results <- data %>% 
        group_by(sample_id, constellation)
    }

    # compute the summary statistics by sample id (and constellation)
    results <- results %>% summarize(average_competence = mean(competence), 
                                     average_relative_competence = mean(relative_competence),
                                     average_accuracy = mean(accurate),
                                     count = n_distinct(sample_id)) %>% 
      # store simulation information
      mutate(population = population, 
             sample = sample, 
             options = options)

    return(results)
  }
}
```

```{r, echo=FALSE}
# test output (set return to data above)
# test <- simulate_single_population(population = 10000, sample = 10, options = 10,
#                                    return = "by degree of majority", distribution = "beta",
#                                    alpha = 2, beta = 6) %>% arrange(sample_id) %>%
#   arrange(sample_id)
# 
# test %>% group_by(relative_majority) %>% summarise(across(c(average_relative_competence,
#                                                         average_accuracy),
#                                                       mean)
#                                                )
```

#### Various populations

```{r, echo=FALSE}
simulate_various_populations <- function(iterations,...) {

  # create data frame with model results for generated samples
  various_populations <- 1:iterations %>% 
    purrr::map_df(function(x){
      # this is essentially a for loop - do the following for each 
      # element in 1:iterations

      results <- simulate_single_population(...)

      # identify iteration
      results$iteration <- x

      # To keep track of progress
      if (x %% 50 == 0) {print(paste("iteration number ", x))}

      return(results)

    }) %>% 
    # store simulation information
    mutate(total_iterations = iterations)

  return(various_populations)
}
```

#### Vary choice options and sample size

This function allows us to investigate how accuracy and competence change with varying the number of choice options and the sample size.

The simulations that this function executes will take quite some time. Therefore, we do not want to run it every time we render this document. Instead we want to store the output of the power simulation in a `.csv` file, and have an integrated "stop" mechanism to prevent execution when that file already exists. To achieve this, we make `file_name` a mandatory argument. If a file with that name already exists, the function will not be executed.

```{r, echo=FALSE}
# set name
file_name <- "model.csv" # change for new analyses / or delete file to re-use same name

vary_sample_options <- function(vary = "options", file_name, n, 
                                format = "result", ...) {

  # only run analysis if a file with that name does not yet exists
  if (!file.exists(paste0("data/", file_name))) {

    # vary choice options for given sample size
    if (vary == "options") {

      # run the populations generating function for different values of n
      data <- n %>% purrr::map_df(function(n_option){
        # this is essentially a for loop - 
        # do the following for each 
        # element of n

        # To keep track of progress
        print(paste("tested option number = ", n_option))

        # run simulation
        result <- simulate_various_populations(
          options = n_option, ...)

        return(result)

      })

      if(format == "result") {
        return(data)
      } else{
        write_csv(data, paste0("data/", file_name))
      }
    }

    # vary sample size for given choice option
    if (vary == "sample") {

      # do the `calculate_power()` function for each sample size and store the results
      # in a new variable called `power`
      data <- n %>% purrr::map_df(function(n_sample){
        # this is essentially a for loop - 
        # do the following for each 
        # element data$n_subj

        # To keep track of progress
        print(paste("tested sample number = ", n_sample))

        # run power calculation
        result <- simulate_various_populations(
          sample = n_sample, ...)

        return(result)

      })

      if(format == "result") {
        return(data)
      } else{
        write_csv(data, paste0("data/", file_name))
      }
    }
  }
}
```

#### Vary competence distribution

We now want to do the above (simulate a single population, repeat this process, do so for various sample sizes and choice options) for different competence distributions. We vary these distributions with the `alpha` and `beta` parameters for the beta distributions.

```{r, echo=FALSE}
# set name
file_name <- "model.csv" # change for new analyses / or delete file to re-use same name

vary_competence <- function(file_name, alpha, beta, names, ...) {

  # only run analysis if a file with that name does not yet exists
  if (!file.exists(paste0("data/", file_name))) {

    competence_data <- tibble(alpha, beta, names) %>% 
      mutate(competence = paste0(names, "\n(alpha = ", alpha, ", beta = ", beta, ")"))

    # do all the above functions for different competence distributions
    data <- competence_data$competence %>% purrr::map_df(function(competence_level){
      # this is essentially a for loop

      # To keep track of progress
      print(paste("competence_level = ", competence_level))

      # set alpha and beta
      alpha = competence_data %>% filter(competence == competence_level) %>% pull(alpha)
      beta = competence_data %>% filter(competence == competence_level) %>% pull(beta)

      # run simulation
      result <- vary_sample_options(
        alpha = alpha, beta = beta, file_name = file_name, ...)

      # add competence level
      result <- result %>% mutate(competence = competence_level, 
                                  alpha = alpha,
                                  beta = beta)

      return(result)

    })

    write_csv(data, paste0("data/", file_name))
  }
}
```

### Plot functions

#### a) Plot various populations

First, we want a function that plots results obtained by the `simulate_various_populations` function. 
```{r, echo=FALSE}
plot_results <- function(data, outcome = "everything") {

  d <- data

  # make constellation a factor with the right levels
  d$constellation <- fct_relevel(d$constellation, "minority(ies)", "dissensus", "majority", "consensus")

  # extract simulation info
  simulation_info <- d %>% summarize(across(c(population, sample, options, total_iterations), mean))

  caption = paste0("Number of choice options: ", simulation_info$options,
                   "; Sample size: ", simulation_info$sample,
                   "; Iterations : ", simulation_info$total_iterations, 
                   "; Population size per iteration: ", simulation_info$population)

  # make descriptive data
  # descriptive <- d %>% 
  #   group_by(constellation) %>% 
  #   summarise(count = sum(count)) %>% 
  #   mutate(rel_freq = round(count / sum(count), digits = 2)) %>% 
  #   gridExtra::tableGrob(rows = NULL)

  # plot for accuracy
  plot_accuracy <- ggplot(d,
                          aes(x = constellation, y = average_accuracy, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = "Convergence", y = "Accuracy") +
    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
    guides(fill = FALSE) +
    plot_theme + 
    theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
    ylim(c(0, 1))

  # plot for competence
  plot_competence <- ggplot(d,
                            aes(x = constellation, y =  average_relative_competence, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = "Convergence", y = "Competence", caption = caption) +
    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
    guides(fill = FALSE) +
    plot_theme + 
    theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
    ylim(c(0, 1))

  # unite tables
  # tables <- gridExtra::grid.arrange(descriptive, simulation_info, ncol = 2)

  if (outcome == "everything") {

    return(plot_accuracy | plot_competence )
  }

  if (outcome == "accuracy") {

    return(plot_accuracy + labs(caption = caption))
  }

  if (outcome == "competence") {

    return(plot_competence)
  }

}
```

#### b) Plot varying sample & options

Second, we want a function that plots results obtained by the `vary_sample_options` function.

```{r, echo=FALSE}
plot_results_vary <- function(data, variable = options) {

  d <- data

  # Extract variable name as a string
  variable_name <- deparse(substitute(variable))

  # retrieve info from choice of variable
  if (variable_name == "options") {
    x_label = "Number of Choice Options"
    title = paste0("Sample: ", d$sample," \n iterations: ", d$total_iterations)
  }
    if (variable_name == "sample") {
    x_label = "Size of informant groups"
    title = paste0("Options: ", d$options," \n iterations: ", d$total_iterations)
  }

  # make constellation a factor with the right levels
  d$constellation <- fct_relevel(d$constellation, "minority(ies)", "dissensus", "majority", "consensus")

  # plot for accuracy
  plot_accuracy <- ggplot(d,
                          aes(x = as.factor({{variable}}), y = average_accuracy, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .5,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = x_label, y = "Accuracy") +
    scale_fill_viridis_d(option = "plasma", begin = 0.1, 
                         limits = rev(levels(d$constellation)),
                         direction = -1
    ) +
    plot_theme

  # plot for competence
  plot_competence <- ggplot(d,
                            aes(x = as.factor({{variable}}), y = average_relative_competence, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .5,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = x_label, y = "Relative competence", fill = NULL) + 
    scale_fill_viridis_d(option = "plasma", begin = 0.1, 
                         limits = rev(levels(d$constellation)),
                         direction = -1
    ) +
    plot_theme

  ggpubr::ggarrange(plot_accuracy, plot_competence, common.legend = TRUE) +
    plot_annotation(title = title)
}
```

#### c) Plot varying competence distributions

Third, we want a function that plots results by various competence levels.

```{r}
plot_competence_vary <- function(data, variable = options, outcome = "Accuracy", 
                                 return = "simulations") {

  d <- data

  # Extract variable name as a string
  variable_name <- deparse(substitute(variable))

  # retrieve info from choice of variable
  if (variable_name == "options") {
    x_label = "Number of Choice Options"
    caption = paste0("Sample size of informants: ", d$sample[1],
                     "; Iterations per sample size: ", d$total_iterations[1], 
                     "; Population size per iteration: ", d$population[1])
  }
  if (variable_name == "sample") {
    x_label = "Size of informant groups"
    caption = paste0("Number of choice options: ", d$options[1],
                     "; Iterations per sample size: ", d$total_iterations[1], 
                     "; Population size per iteration: ", d$population[1])
  }

  # make constellation a factor with the right levels
  d$constellation <- fct_relevel(d$constellation, "minority(ies)", "dissensus", "majority", "consensus")

  d <- d %>% 
    group_by(constellation, {{variable}}, competence) %>% 
    summarise(competence_mean = mean(average_relative_competence),
              accuracy_mean = mean(average_accuracy),
              accuracy_sd = sd(average_accuracy),
              competence_sd = sd(average_relative_competence)
    )

  # Assign outcome variable
  if(outcome == "Accuracy") {
    d$outcome <- d$accuracy_mean
    d$outcome_sd <- d$accuracy_sd
  }

  if(outcome == "Competence") {
    d$outcome <- d$competence_mean
    d$outcome_sd <- d$competence_sd
  }

  # plot for accuracy
  plot_outcome <- ggplot(d,
                         aes(x = {{variable}}, y = outcome,
                             color = constellation)) +
    # geom_half_violin(position = position_nudge(x = -.2),
    #                  adjust=2, alpha = .5,
    #                  side = "l") +
    geom_pointrange(aes(ymin = outcome - outcome_sd, 
                        ymax = outcome + outcome_sd), 
                    size = 0.1) +
    geom_line() +
    # Add nice labels
    labs(x = x_label, y = outcome, caption = caption) +
    scale_color_viridis_d(option = "plasma", begin = 0.1, 
                          limits = rev(levels(d$constellation)),
                          direction = -1
    ) +
    facet_wrap(~competence) +
    plot_theme + 
    theme(legend.position = "top")

  # Plot competence distributions
  # Your data frame
  d <- data %>% 
    group_by(competence) %>% 
    reframe(alpha = unique(alpha), beta = unique(beta))

  # Function to plot the different competence distributions 
  plot_competence_distributions <- function(alpha, beta) {
    ggplot() +
      geom_function(
        fun = ~dbeta(.x, shape1 = alpha, shape2 = beta)
      ) +
      labs(title = paste0("alpha = ", alpha, ", beta = ", beta), 
           y = "Density", x = "Competence") +
      plot_theme
  }

  # Specify the values of alpha and beta from the data
  alpha_values <- d$alpha
  beta_values <- d$beta

  # Create a list of ggplot objects
  plot_list <- map2(alpha_values, beta_values, plot_competence_distributions)

  # Arrange and display the plots
  competence_plot <- ggarrange(plotlist = plot_list) %>%
    ggpubr::annotate_figure(top = ggpubr::text_grob("Competence distributions", 
                                                    face = "bold",
                                                    color = "darkgrey",
                                                    size = 14))

  if(return == "competence distributions") {
    return(competence_plot)
  } else {
   return(plot_outcome) 
  }
} 
```

#### d) Plot varying competence distributions by relative majority

Fourth, we want another function that plots results by various competence levels - but this time not for  constellations but the numeric value of the relative majority.

```{r}
plot_competence_vary_relative_majority <- function(data, variable = options, outcome = "Accuracy") {

  d <- data

  # Extract variable name as a string
  variable_name <- deparse(substitute(variable))

  # retrieve info from choice of variable
  if (variable_name == "options") {
    caption = paste0("Sample size of informants: ", d$sample[1],
                     "; Iterations per sample size: ", d$total_iterations[1], 
                     "; Population size per iteration: ", d$population[1])
  }
  if (variable_name == "sample") {
    caption = paste0("Number of choice options: ", d$options[1],
                     "; Iterations per sample size: ", d$total_iterations[1], 
                     "; Population size per iteration: ", d$population[1])
  }

  d <- d %>% 
    group_by(relative_majority, {{variable}}, competence) %>%
    summarise(competence_mean = mean(average_relative_competence),
              accuracy_mean = mean(average_accuracy),
              accuracy_sd = sd(average_accuracy),
              competence_sd = sd(average_relative_competence)
    )

  # Assign outcome variable
  if(outcome == "Accuracy") {
    d$outcome <- d$accuracy_mean
    d$outcome_sd <- d$accuracy_sd
  }

  if(outcome == "Competence") {
    d$outcome <- d$competence_mean
    d$outcome_sd <- d$competence_sd
  }

  plot_outcome <- ggplot(d, 
                         aes(x = relative_majority, y = outcome,
                             color = as.factor({{variable}}))) +
    geom_point() +
    geom_line() +
    # Add nice labels
    labs(x = "Relative majority \n (share of informants agreeing on an option)", 
         y = outcome, caption = caption, 
         color = variable_name) +
    scale_color_viridis_d(option = "plasma", begin = 0.1, 
                          #limits = rev(levels(d$constellation)),
                          direction = -1
    ) +
    facet_wrap(~competence) +
    plot_theme + 
    theme(legend.position = "top")

   return(plot_outcome)

} 
```

#### e) Plot competence

Finally, we want to plot the underlying competence distributions

```{r}
# Plot competence distributions
plot_competence_distributions <- function(data) {
  # Your data frame
  d <- data %>% 
    group_by(competence) %>% 
    reframe(alpha = unique(alpha), beta = unique(beta))

  # Function to plot the different competence distributions 
  plot_competence_distributions <- function(alpha, beta) {
    ggplot() +
      geom_function(
        fun = ~dbeta(.x, shape1 = alpha, shape2 = beta)
      ) +
      labs(title = paste0("alpha = ", alpha, ", beta = ", beta), 
           y = "Density", x = "Competence") +
      plot_theme
  }

  # Specify the values of alpha and beta from the data
  alpha_values <- d$alpha
  beta_values <- d$beta

  # Create a list of ggplot objects
  plot_list <- map2(alpha_values, beta_values, plot_competence_distributions)

  # Arrange and display the plots
  competence_plot <- ggarrange(plotlist = plot_list) %>%
    ggpubr::annotate_figure(top = ggpubr::text_grob("Competence distributions", 
                                                    face = "bold",
                                                    color = "darkgrey",
                                                    size = 14))

  return(competence_plot) 
}
```

## Simulation

```{r}
# Tryout beta functions
# ggplot() +
#   geom_function(fun = ~dbeta(.x, shape1 = 0.9, shape2 = 0.9))
```

### Experiment 4

#### Generate Data

First, we want to simulate data under assumptions that correspond to our experimental design.

For the first, simulation, we fix all parameters. We assume:
- a uniform competence distribution (beta = 1 and alpha = 1)
- 3 choice options,
- samples of 3 (i.e. 3 informants)

We generate a sample of 999x1000 = 999000 individuals.

```{r, message=FALSE}
file_name <- "sim_experiment_4.csv" # change for new analyses / or delete file to re-use same name

if (!file.exists(paste0("data/", file_name))) {

  # generate data
  data <- simulate_various_populations(iterations = 1000, population = 999, sample = 3, options = 3)

  write_csv(data, paste0("data/", file_name))
}

# read simulated data from .csv files
sim_experiment_4 <- read_csv(paste0("data/", file_name))
```

#### Plot result

We make then make a plot corresponding to the figure from the experiments.

```{r}
# descriptive
sim_experiment_4 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))

plot_results(sim_experiment_4)
```

#### Analyze data

```{r, echo=FALSE}
# possible analysis
regression_data <- sim_experiment_4 %>%
  mutate(
    # make numeric version of convergence
    convergence = case_when( 
      constellation == "minority" ~ 0, 
      constellation == "dissensus" ~ 1,
      constellation == "majority" ~ 2,
      constellation == "consensus" ~ 3), 
    # scale accuracy to reach from 0 to 100 (instead from 0 to 1)
    average_accuracy = average_accuracy*100,
    # scale competence to reach from 1 to 7 (instead of 0 to 1)
    average_relative_competence = 6*average_relative_competence + 1 
  )

accuracy_model <- lm(average_accuracy ~ convergence, 
                     data = regression_data)
competence_model <- lm(average_relative_competence ~ convergence, 
                       data = regression_data)
```

#### Compare participants and simulations

```{r, message=FALSE}
# read data from experiments
d <- read_csv("./data/cleaned.csv")

# make a categorical variable from `convergence`
d <- d %>% 
  mutate(convergence_categorical = recode_factor(convergence, 
                                                 `0` = "opposing majority", 
                                                 `1` = "divergence", 
                                                 `2` = "majority", 
                                                 `3` = "consensus",
                                                 .default = NA_character_)
         )
```

##### Table

```{r, echo=FALSE}
# Calculate models for participants

# model for accuracy
# random intercept and slope by participants
accuracy_participants <- lmer(accuracy ~ convergence + (1 + convergence | id), 
                                 data = d)
# model for competence
# random intercept and slope by participants
competence_participants <- lmer(competence ~ convergence + 
                           (1 + convergence | id), data = d)

# main result table
modelsummary::modelsummary(list("Accuracy" = accuracy_participants, 
                                "Competence" = competence_participants,
                                "Accuracy" = accuracy_model, 
                                "Competence" = competence_model
                                ),
                           title = 'Participants vs. Model', 
                           stars = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "Participants" = 2, "Model" = 2))
```

##### Plots

```{r, echo=FALSE}
# Participant data

# plot for accuracy
plot_accuracy <- ggplot(d,
       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# plot for competence
plot_competence <- ggplot(d,
       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r}
# Model data
plot_accuracy_model <- plot_results(sim_experiment_4, outcome = "accuracy")
plot_competence_model <- plot_results(sim_experiment_4, outcome = "competence")
```

```{r, echo=FALSE}
# Combine participant and model
# accuracy
plot_accuracy <- plot_accuracy + ggtitle("Participants")
plot_accuracy_model <- plot_accuracy_model + ggtitle("Model")

plot_accuracy + plot_accuracy_model
```

```{r, echo=FALSE}
# competence
plot_competence <- plot_competence + ggtitle("Participants")
plot_competence_model <- plot_competence_model + ggtitle("Model")

plot_competence + plot_competence_model
```

### Experiment 6

#### Generate data

In experiment 6, we varied choice options (from 3 to 10).

In this simulation, we assume:
- a sample size of 3 
- a uniform competence distribution (beta = 1 and alpha = 1)

```{r, message=FALSE}
n <- c(3, 5, 10, 20, 50, 100)

# run simulation and store results in .csv files
vary_sample_options(n = n, format = "csv file", vary = "options", iterations = 5, population = 10000, 
                    sample = 3, file_name = "sim_vary_options_3_sample.csv", 
                    distribution = "beta", alpha = 1, beta = 1)

# read simulated data from .csv files
data_3_options <- read_csv("data/sim_vary_options_3_sample.csv")
```

#### Plot result

```{r}
# example for plot
plot_results_vary(data_3_options, 
                  variable = options)
```

#### Analyze data

```{r, echo=FALSE}
# possible analysis for varying choice options (3 vs. 10)
regression_data <- data_3_options %>%
  filter(options %in% c(3, 10)) %>%
  # group_by(iteration, constellation, options) %>%
  # summarise(across(c(average_accuracy, average_relative_competence), mean)) %>%
  mutate(
    convergence = case_when(
      constellation == "minority" ~ 0,
      constellation == "dissensus" ~ 1,
      constellation == "majority" ~ 2,
      constellation == "consensus" ~ 3),
    options = as.factor(options),
    # scale accuracy to reach from 0 to 100 (instead from 0 to 1)
    average_accuracy = average_accuracy*100,
    # scale competence to reach from 1 to 7 (instead of 0 to 1)
    average_relative_competence = 6*average_relative_competence + 1
  )

accuracy <- lm(average_accuracy ~ convergence + options + convergence*options,
               data = regression_data)
competence <- lm(average_relative_competence ~ convergence + options + convergence*options, 
                 data = regression_data)

summary(competence)
```

### Vary competence

Finally, we want to vary the underlying competence distribution. Along with it, we vary either sample size or number of choice options (holding the respective other constant).

#### By constellation

We classify samples into categories: dissensus, minority (any of potentially several minorites), majority (the answer with the most counts), and consensus.

```{r, message=FALSE}
n <- c(3, 5, 10, 20, 50, 100)
alpha = c(0.1, 2, 1, 1, 6, 0.9)
beta = c(1, 6, 1, 0.1, 2, 0.9)
names = c("very incompetent", "incompetent", "uniform", 
          "very competent", "competent", "bimodal")

vary_competence(file_name = "sim_vary_competence_sample_3_options.csv", 
                distribution = "beta", 
                alpha = alpha, beta = beta, names = names, vary = "sample", n = n, 
                iterations = 5, 
                population = 1000, options = 3)

varied_competence <- read_csv("data/sim_vary_competence_sample_3_options.csv")
```

```{r}
plot_competence_vary(varied_competence,  return = "competence distributions")
plot_competence_vary(varied_competence , outcome = "Competence", variable = sample)
plot_competence_vary(varied_competence , outcome = "Accuracy", variable = sample)
```
#### By relative majority

Here, we do not group data into categories, but simply sort it by the share of participants (within a sample of informants) that agreed with an options. We call this relative majority.

##### Vary sample size

```{r}
n <- c(3, 5, 10, 20, 50, 100)
alpha = c(0.1, 2, 1, 1, 6, 0.9)
beta = c(1, 6, 1, 0.1, 2, 0.9)
names = c("very incompetent", "incompetent", "uniform", 
          "very competent", "competent", "bimodal")

# Vary the sample size, while holding constant on 3 choice options
vary_competence(file_name = "sim_rel_majority_vary_competence_sample_3_options.csv", 
                distribution = "beta", 
                alpha = alpha, beta = beta, names = names, vary = "sample", n = n, 
                iterations = 100, 
                population = 1000, options = 3, return = "by degree of majority")

vary_competence_sample <- read_csv("data/sim_rel_majority_vary_competence_sample_3_options.csv")
```

```{r}
# Competence distributions
plot_competence_distributions(vary_competence_sample)
```

```{r}
plot_competence_vary_relative_majority(vary_competence_sample,
                                       outcome = "Accuracy", variable = sample)
```
```{r}
plot_competence_vary_relative_majority(vary_competence_sample,
                                       outcome = "Competence", variable = sample)
```

##### Vary choice options

```{r}
n <- c(3, 5, 10, 20, 50, 100)
alpha = c(0.1, 2, 1, 1, 6, 0.9)
beta = c(1, 6, 1, 0.1, 2, 0.9)
names = c("very incompetent", "incompetent", "uniform", 
          "very competent", "competent", "bimodal")

# Vary the sample size, while holding constant on 3 choice options
vary_competence(file_name = "sim_rel_majority_vary_competence_options_3_sample.csv", 
                distribution = "beta", 
                alpha = alpha, beta = beta, names = names, vary = "options", n = n, 
                iterations = 100, 
                population = 1000, sample = 3, return = "by degree of majority")

vary_competence_options <- read_csv("data/sim_rel_majority_vary_competence_options_3_sample.csv")
```

```{r}
# Competence distributions
plot_competence_distributions(vary_competence_options)
```

```{r}
plot_competence_vary_relative_majority(vary_competence_options,
                                       outcome = "Accuracy", variable = options)
```

```{r}
plot_competence_vary_relative_majority(vary_competence_options,
                                       outcome = "Competence", variable = options)
```