### 2024-01-18 appendix_exp2.Rmd

# Experiment 2 {#exp2}

## Design

\FloatBarrier

We manipulated informational dependency. According to the condition, players read a different introduction before seeing a set of estimates (see Table \@ref(tab:exp2-conditions)).

```{r exp2-conditions}
# The different introductions that participants read before seeing the sets of estimates, according to the conditions 
data.frame(
  Condition = c("Independence", "Discussion"),
  Description = c(
    "Players are asked to make completely independent decisions – they cannot see each other’s estimates, or talk with each other before giving their estimates.",
    "Players are asked to talk with each other about the game at length before giving their estimates."
  )
) #%>% 
  # apa_table(align = c("m{3cm}", "m{12cm}")) 
```

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

## Results

Figure \@ref(fig:exp2-plot) visualizes the results and table \@ref(tab:exp2-table) contains descriptive results.

```{r exp2-table}
exp2 %>% 
  group_by(independence) %>% 
  summarise(across(c(competence, confidence), list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(confidence_mean, " (sd = ", confidence_sd, ")"),
         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")")
         ) %>% 
  rename(Independence = independence) %>% 
  select(Independence, Accuracy, Competence) %>% 
  apa_table()
```

(ref:exp2-plot) Distributions of accuracy and competence by level of informational dependency.

```{r exp2-plot, fig.cap="(ref:exp2-plot)"}
# make mean data for plot
x_nudge <- 0.8

means <- exp2 %>%
  group_by(independence) %>%
  summarize(confidence = mean(confidence), 
            competence = mean(competence)) %>% 
  pivot_longer(c(confidence, competence), 
               names_to = "outcome", 
               values_to = "value") %>% 
  mutate(x_position = ifelse(independence == "dependent", value + x_nudge, 
                                value - x_nudge))

confidence_plot <- ggplot(data=exp2, aes(x=confidence, fill=independence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
 # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "confidence"), 
             aes(xintercept = value, 
                 color = independence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "confidence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1) +
  scale_color_brewer(palette = "Accent",
                    direction = 1) +
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

competence_plot <- ggplot(data=exp2, aes(x=competence, fill=independence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
 # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "competence"), 
             aes(xintercept = value, 
                 color = independence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "competence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1) +
  scale_color_brewer(palette = "Accent",
                    direction = 1) +
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

figure <- ggarrange(confidence_plot, 
                    competence_plot + 
                      theme(axis.text.y = element_blank(),
                            axis.ticks.y = element_blank(),
                            axis.title.y = element_blank()), 
                    common.legend = T) +
  theme(legend.position="top") 
figure
```