### 2024-01-18 appendix_exp4.Rmd

# Experiment 4 {#exp4}

\FloatBarrier

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

## Stimuli

We manipulate convergence within participants. All participants see all four conditions, with two stimuli (i.e. game results) per condition (see Table \@ref(tab:exp4-stimuli)). Each participant therefore sees eight stimuli in total (4 convergence levels x 2 stimuli)

```{r exp4-stimuli, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(condition = c("opposing majority (0)", 
                                        "dissensus (1)", "majority (2)", 
                                        "consensus (3)"),
                          imgage_a = c("![](Experiment_4/figures/stimuli/opp_majority_a.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/divergence_a.png){ width=60% }",
                        "![](Experiment_4/figures/stimuli/majority_a.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/consensus_a.png){ width=60% }"),
                        imgage_b = c("![](Experiment_4/figures/stimuli/opp_majority_b.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/divergence_b.png){ width=60% }",
                        "![](Experiment_4/figures/stimuli/majority_b.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/consensus_b.png){ width=60% }"))

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Level", "Version a)", "Version b)"), 
                  align = "c",
                  caption = "All stimuli by levels of convergence")

```

## Results

Table \@ref(tab:exp4-table) contains descriptive results.

```{r exp4-table}
exp4 %>% 
  group_by(convergence_categorical, convergence) %>% 
  summarise(across(c(competence, accuracy), list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ")"),
         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")"),
         Convergence = paste0(convergence_categorical, " (", convergence, ")")
         ) %>% 
  ungroup() %>% 
  select(Convergence, Accuracy, Competence) %>% 
  apa_table()
```