### 2024-03-26 appendix_exp4.Rmd

# Experiment 4 {#exp4}

\FloatBarrier

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

## Stimuli

We manipulate convergence within participants. All participants see all four conditions, with two stimuli (i.e. game results) per condition (see Table \@ref(tab:exp4-stimuli)). Each participant therefore sees eight stimuli in total (4 convergence levels x 2 stimuli)

```{r exp4-stimuli, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(condition = c("minority (0)", 
                                        "dissensus (1)", "majority (2)", 
                                        "consensus (3)"),
                          imgage_a = c("![](Experiment_4/figures/stimuli/opp_majority_a.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/divergence_a.png){ width=60% }",
                        "![](Experiment_4/figures/stimuli/majority_a.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/consensus_a.png){ width=60% }"),
                        imgage_b = c("![](Experiment_4/figures/stimuli/opp_majority_b.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/divergence_b.png){ width=60% }",
                        "![](Experiment_4/figures/stimuli/majority_b.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/consensus_b.png){ width=60% }"))

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Level", "Version a)", "Version b)"), 
                  align = "c",
                  caption = "All stimuli by levels of convergence") %>%
  column_spec(1, width = "2em") %>%
  column_spec(2, width = "15em") %>%
  column_spec(3, width = "15em")

```

## Results

Table \@ref(tab:exp4-table) contains descriptive results and Figure \@ref(fig:exp4-plot) visualizes the results.

```{r exp4-table}
exp4 %>% 
  group_by(convergence_categorical, convergence) %>% 
  summarise(across(c(competence, accuracy), list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ")"),
         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")"),
         Convergence = paste0(convergence_categorical, " (", convergence, ")")
         ) %>% 
  ungroup() %>% 
  select(Convergence, Accuracy, Competence) %>% 
  apa_table()
```

(ref:exp4-plot) Distributions of accuracy and competence by level of convergence.

```{r exp4-plot, fig.cap="(ref:exp4-plot)"}
# plot for accuracy
plot_accuracy <- ggplot(exp4,
       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +  
  rremove("xlab")   # Remove x-axis label

# plot for competence
plot_competence <- ggplot(exp4,
       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  scale_y_continuous(limits = c(1,7), breaks = seq(1,7)) +
  labs(x = "Convergence", y = "Competence") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +  
  rremove("xlab")   # Remove x-axis label

ggarrange(plot_accuracy, plot_competence) %>% 
  annotate_figure(bottom = textGrob("Convergence",gp = gpar(cex = 1, fontface = "bold")))
```

## Comparison with simulation

We compared the accuracy and competence ratings of the participants to actual accuracy and competence values from simulated data. The data was generated using the model described in the main paper, and assuming a uniform competence distribution. Compared to this data generating model, participants underestimate the effect of convergence for both accuracy and competence, but more for accuracy. Table \@ref(tab:model-vs-participant-table) compares the results of regressions on participant data and regressions on simulated data. Fig. \@ref(fig:exp4-model) visualizes the results.

```{r}
# read simulated data from .csv files
sim_experiment_4 <- read_csv("Experiment_4/data/sim_experiment_4.csv")
```

```{r}
# Run model on simulated data
regression_data <- sim_experiment_4 %>%
  mutate(
    # make numeric version of convergence
    convergence = case_when( 
      constellation == "minority" ~ 0, 
      constellation == "dissensus" ~ 1,
      constellation == "majority" ~ 2,
      constellation == "consensus" ~ 3), 
    # scale accuracy to reach from 0 to 100 (instead from 0 to 1)
    average_accuracy = average_accuracy*100,
    # scale competence to reach from 1 to 7 (instead of 0 to 1)
    average_relative_competence = 6*average_relative_competence + 1 
  )

accuracy_model <- lm(average_accuracy ~ convergence, 
                     data = regression_data)
competence_model <- lm(average_relative_competence ~ convergence, 
                       data = regression_data)
```

```{r}
# read participant data
d <- exp4
```

```{r model-vs-participant-table}
# Calculate models for participants

# model for accuracy
# random intercept and slope by participants
accuracy_participants <- lmer(accuracy ~ convergence + (1 + convergence | id), 
                                 data = d)
# model for competence
# random intercept and slope by participants
competence_participants <- lmer(competence ~ convergence + 
                           (1 + convergence | id), data = d)

# main result table
modelsummary::modelsummary(list("Accuracy" = accuracy_participants, 
                                "Competence" = competence_participants,
                                "Accuracy" = accuracy_model, 
                                "Competence" = competence_model
                                ),
                           title = 'Participants vs. Model (Exp. 4)', 
                           stars = TRUE) %>%
  kableExtra::add_header_above(c(" " = 1, "Participants" = 2, "Model" = 2))
```

```{r, echo=FALSE}
# Participant data

# plot for accuracy
plot_accuracy <- ggplot(d,
       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# plot for competence
plot_competence <- ggplot(d,
       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r}
# Model data
plot_accuracy_model <- plot_results(sim_experiment_4, outcome = "accuracy")
plot_competence_model <- plot_results(sim_experiment_4, outcome = "competence")
```

(ref:exp4-model) Comparison of accuracy (top) and competence (bottom) ratings in participant data (left) to observed accuracy in simulated data (right), by level of convergence.

```{r exp4-model, fig.cap="(ref:exp4-model)"}
# Combine participant and model
# accuracy
plot_accuracy <- plot_accuracy + ggtitle("Participants") + labs(caption = NULL)
plot_accuracy_model <- plot_accuracy_model + ggtitle("Model") + labs(caption = NULL)

# competence
plot_competence <- plot_competence 
plot_competence_model <- plot_competence_model

(plot_accuracy + plot_accuracy_model) / (plot_competence + plot_competence_model)
```