### 2024-03-26 appendix_exp5.Rmd

# Experiment 5 {#exp5}

## Design

\FloatBarrier

We manipulated convergence within participants in the same way we did in experiment 4. In addition, between participants, we manipulated informational dependence, akin to experiment 3 (Table \@ref(tab:exp5-condition)). In the biased condition, experts were described to gain personally from recommending a certain investment option - but without specifying what that option is. In the independent condition, there was no such conflict of interest and experts were described as independent.

```{r exp5-condition}
# Participants read only one of these introductions before seeing the sets of estimates, depending on which condition they had been assigned to
data.frame(
  Condition = c("Independence condition", "Conflict of interest condition"),
  Description = c(
    "The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.",
    "The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations."
  )
) %>% 
  apa_table(align = c("m{3cm}", "m{12cm}")) 
```

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

## Results

Table \@ref(tab:exp5-table) contains descriptive results.

```{r exp5-table}
exp5_accuracy <- exp5 %>% 
  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
  group_by(independence, Convergence, convergence) %>% 
  summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
  select(independence, Accuracy, Convergence, convergence) %>% 
mutate(independence = ifelse(independence == "conflict of interest", "Conflict of interest", 
                "Independent") 
         ) %>% 
  pivot_wider(names_from = independence,
              values_from = Accuracy) %>% 
  arrange(convergence) %>% 
  ungroup() %>% 
  select(-convergence) %>% 
  rename(" " = Convergence) %>% 
  as.matrix()

exp5_competence <- exp5 %>% 
  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
  group_by(independence, Convergence, convergence) %>% 
  summarise(across(competence, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
  select(independence, Competence, Convergence, convergence) %>% 
mutate(independence = ifelse(independence == "conflict of interest", "Conflict of interest", 
                "Independent") 
         ) %>% 
  pivot_wider(names_from = independence,
              values_from = Competence) %>% 
  ungroup() %>% 
  arrange(convergence) %>% 
  select(-c(convergence, Convergence)) %>% 
  as.matrix()

table <- cbind(exp5_accuracy, exp5_competence)

table %>%
  kable(booktabs = TRUE, 
        # need to add a (empty) caption for cross referencing to work
        caption = "") %>%
  kable_styling() %>% 
  add_header_above(c("", "Accuracy" = 2, "Competence" = 2)) %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down") 
```